---
title: "Week 4 - Homework"
author: "STAT 420, Summer 2020, Chenyue Jiao(cjiao4)"
date: ''
output:
  html_document: 
    theme: readable
    toc: yes  
  pdf_document: default
urlcolor: cyan
---

***

## Exercise 1 (Using `lm`)

For this exercise we will use the data stored in [`nutrition-2018.csv`](nutrition-2018.csv). It contains the nutritional values per serving size for a large variety of foods as calculated by the USDA in 2018. It is a cleaned version totaling 5956 observations and is current as of April 2018.

The variables in the dataset are:

- `ID` 
- `Desc` - short description of food
- `Water` - in grams
- `Calories` 
- `Protein` - in grams
- `Fat` - in grams
- `Carbs` - carbohydrates, in grams
- `Fiber` - in grams
- `Sugar` - in grams
- `Calcium` - in milligrams
- `Potassium` - in milligrams
- `Sodium` - in milligrams
- `VitaminC` - vitamin C, in milligrams
- `Chol` - cholesterol, in milligrams
- `Portion` - description of standard serving size used in analysis

**(a)** Fit the following multiple linear regression model in `R`. Use `Calories` as the response and `Fat`, `Sugar`, and `Sodium` as predictors.

\[
Y_i = \beta_0 + \beta_1 x_{i1} + \beta_2 x_{i2} + \beta_3 x_{i3} + \epsilon_i.
\]

Here,

- $Y_i$ is `Calories`.
- $x_{i1}$ is `Fat`.
- $x_{i2}$ is `Sugar`.
- $x_{i3}$ is `Sodium`.

Use an $F$-test to test the significance of the regression. Report the following:
 
- The null and alternative hypotheses
- The value of the test statistic
- The p-value of the test
- A statistical decision at $\alpha = 0.01$
- A conclusion in the context of the problem

When reporting these, you should explicitly state them in your document, not assume that a reader will find and interpret them from a large block of `R` output.

```{r message=FALSE}
library(readr)
nutrition = read_csv("nutrition-2018.csv")
null_nutrition_model = lm(Calories ~ 1, data = nutrition)
full_nutrition_model = lm(Calories ~ Fat + Sugar + Sodium, data = nutrition)
anova(null_nutrition_model, full_nutrition_model)
summary(full_nutrition_model)$fstatistic
```

- The null and alternative hypotheses: 
- The null hypothesis: $H_0: \beta_1 = \beta_2 = \beta_3 = 0$;
- The alternative hypothesis: $H_1$: At least one of $\beta_j \neq 0, j=1,2,...,(p-1)$.

- The value of the test statistic: 6590.94

- The p-value of the test: 0 (< 2.2e-16)

- A statistical decision at $\alpha = 0.01$
- The value of the F statistic is 6590.94, and the p-value is extremely low, so we reject the null hypothesis at $\alpha = 0.01$ and say that the regression is significant. 

- A conclusion in the context of the problem
- At least one of fat, sugar or sodium has a significant linear relationship with calories.


**(b)** Output only the estimated regression coefficients. Interpret all $\hat{\beta}_j$ coefficients in the context of the problem.

```{r}
coef(full_nutrition_model)
```

- $\hat\beta_0$ = 1.004561e+02 is the estimate for $\beta_0$, the mean calories per serving size for a food that contains 0 gram fat, 0 gram sugar and 0 milligram sodium.
- $\hat\beta_1$ = 8.483289e+00 is the estimate for $\beta_1$, the average change in calories per serving size for an increase in fat of one grams for a food that contains a certain grams sugar and a certain milligrams sodium.
- $\hat\beta_2$ = 3.900517e+00 is the estimate for $\beta_2$, the average change in calories per serving size for an increase in sugar of one grams for a food that contains a certain grams fat and a certain milligrams sodium.
- $\hat\beta_3$ = 6.165246e-03 is the estimate for $\beta_3$, the average change in calories per serving size for an increase in sodium of one milligrams for a food that contains a certain grams fat and a certain grams sugar.

**(c)** Use your model to predict the number of `Calories` in a Big Mac. According to [McDonald's publicized nutrition facts](https://www.mcdonalds.com/us/en-us/about-our-food/nutrition-calculator.html), the Big Mac contains 30g of fat, 9g of sugar, and 1010mg of sodium.

```{r}
predict(full_nutrition_model, newdata = data.frame(Fat = 30, Sugar = 9, Sodium = 1010))
```

**(d)** Calculate the standard deviation, $s_y$, for the observed values in the Calories variable. Report the value of $s_e$ from your multiple regression model. Interpret both estimates in the context of this problem.

```{r}
s_y = sd(nutrition$Calories)
s_y
s_e = summary(full_nutrition_model)$sigma
s_e
```

- $s_y = 168.05$ is the estimate of the variability of calories, that is, how the observed calories data varies about its mean. 
- $s_e = 80.8543$ is the estimate of the variability of the residuals of the model, that is, how the observed calories data varies about the fitted model(fat, sugar, and sodium). 

**(e)** Report the value of $R^2$ for the model. Interpret its meaning in the context of the problem.

```{r}
summary(full_nutrition_model)$r.squared
```

We could say that 76.86281% for the observed variation in calories is explained by the linear relationship with the three predictor variables, fat, sugar and sodium.

**(f)** Calculate a 90% confidence interval for $\beta_2$. Give an interpretation of the interval in the context of the problem.

```{r}
confint(full_nutrition_model, level = 0.90)
confint(full_nutrition_model, level = 0.90)[3, ]
```

We are 90% confident that the average change in calories per serving size for an increase in sugar of one grams is between 3.783051 and 4.017983.

**(g)** Calculate a 95% confidence interval for $\beta_0$. Give an interpretation of the interval in the context of the problem.

```{r}
confint(full_nutrition_model, level = 0.95)
confint(full_nutrition_model, level = 0.95)[1, ]
```

We are 95% confident that the mean calories per serving size for a food that contains 0 gram fat, 0 gram sugar and 0 milligram sodium is between 97.69443 and 103.21768.

**(h)** Use a 99% confidence interval to estimate the mean Calorie content of a food with 23g of fat, 0g of sugar, and 400mg of sodium, which is true of a large order of McDonald's french fries. Interpret the interval in context.
 
```{r}
predict(full_nutrition_model, newdata = data.frame(Fat = 23, Sugar = 0, Sodium = 400), interval = "confidence", level = 0.99)
```

We are 99% confident that the mean calories per serving size for the large order of McDonald's french fries that contains 23 grams fat, 0 gram sugar and 400 milligrams sodium is between 294.3532 and 301.7224.
 
**(i)** Use a 99% prediction interval to predict the Calorie content of a Crunchwrap Supreme, which has 21g of fat, 6g of sugar, and 1200mg of sodium according to [Taco Bell's publicized nutrition information](https://www.tacobell.com/nutrition/info). Interpret the interval in context.

```{r}
predict(full_nutrition_model, newdata = data.frame(Fat = 21, Sugar = 6, Sodium = 1200), interval = "prediction", level = 0.99)
```

We are 99% confident that the calories per serving size for the Crunchwrap Supreme that contains 21 grams fat, 6 grams sugar and 1200 milligrams sodium is between 101.0345 and 517.7786.

***

## Exercise 2 (More `lm` for Multiple Regression)

For this exercise we will use the data stored in [`goalies.csv`](goalies.csv). It contains career data for 462 players in the National Hockey League who played goaltender at some point up to and including the 2014-2015 season. The variables in the dataset are:
 
- `W` - Wins
- `GA` - Goals Against
- `SA` - Shots Against
- `SV` - Saves
- `SV_PCT` - Save Percentage
- `GAA` - Goals Against Average
- `SO` - Shutouts
- `MIN` - Minutes
- `PIM` - Penalties in Minutes

For this exercise we will consider three models, each with Wins as the response. The predictors for these models are:

- Model 1: Goals Against, Saves
- Model 2: Goals Against, Saves, Shots Against, Minutes, Shutouts
- Model 3: All Available

**(a)** Use an $F$-test to compares Models 1 and 2. Report the following:

- The null hypothesis
- The value of the test statistic
- The p-value of the test
- A statistical decision at $\alpha = 0.05$
- The model you prefer

```{r message=FALSE}
library(readr)
goalies = read_csv("goalies.csv")
```

```{r}
goalies_model_1 = lm(W ~ GA + SV, data = goalies)
goalies_model_2 = lm(W ~ GA + SV + SA + MIN + SO, data = goalies)
anova(goalies_model_1, goalies_model_2)
```

- The null hypothesis
- The null hypothesis is $H_0: \beta_{\texttt{SA}} = \beta_{\texttt{MIN}} = \beta_{\texttt{SO}} = 0$.

- The value of the test statistic: 462.59
- The p-value of the test: 0 (< 2.2e-16)

- A statistical decision at $\alpha = 0.05$
- The value of the F statistic is 462.59, and the p-value is extremely low, so we reject the null hypothesis at $\alpha = 0.05$.

- A conclusion in the context of the problem
- At least one of Shots Against, Minutes, Shutouts has a significant linear relationship with Goals Against, Saves already in the model. 

- The model you prefer
- I prefer the model 2 with the predictors of Goals Against, Saves, Shots Against, Minutes, Shutouts. 

**(b)** Use an $F$-test to compare Model 3 to your preferred model from part **(a)**. Report the following:

- The null hypothesis
- The value of the test statistic
- The p-value of the test
- A statistical decision at $\alpha = 0.05$
- The model you prefer

```{r}
goalies_model_3 = lm(W ~ .,, data = goalies)
anova(goalies_model_2, goalies_model_3)
```

- The null hypothesis
- The null hypothesis is $H_0: \beta_{\texttt{SV_PCT}} = \beta_{\texttt{GAA}} = \beta_{\texttt{PIM}} = 0$.

- The value of the test statistic: 4.052
- The p-value of the test: 0.007353

- A statistical decision at $\alpha = 0.05$
- The value of the F statistic is 4.052, and the p-value is extremely low, so we reject the null hypothesis at $\alpha = 0.05$.

- The model you prefer
- I prefer the model 3 with all predictors, because at least one of SV_PCT, GAA, PIM has a significant linear relationship with the predictors already in model 2.

**(c)** Use a $t$-test to test $H_0: \beta_{\texttt{SV}} = 0 \ \text{vs} \ H_1: \beta_{\texttt{SV}} \neq 0$ for the model you preferred in part **(b)**. Report the following:

- The value of the test statistic
- The p-value of the test
- A statistical decision at $\alpha = 0.05$

```{r}
summary(goalies_model_3)$coefficients
```

- The value of the test statistic: -3.8577392
- The p-value of the test: 1.310371e-04

- A statistical decision at $\alpha = 0.05$
- The value of the test statistic is -3.8577392, and the p-value is extremely low, so we reject the $H_0$ hypothesis at $\alpha = 0.05$.

***

## Exercise 3 (Regression without `lm`)

For this exercise we will once again use the `Ozone` data from the `mlbench` package. The goal of this exercise is to fit a model with `ozone` as the response and the remaining variables as predictors.

```{r}
data(Ozone, package = "mlbench")
Ozone = Ozone[, c(4, 6, 7, 8)]
colnames(Ozone) = c("ozone", "wind", "humidity", "temp")
Ozone = Ozone[complete.cases(Ozone), ]
```

**(a)** Obtain the estimated regression coefficients **without** the use of `lm()` or any other built-in functions for regression. That is, you should use only matrix operations. Store the results in a vector `beta_hat_no_lm`. To ensure this is a vector, you may need to use `as.vector()`. Return this vector as well as the results of `sum(beta_hat_no_lm ^ 2)`.

```{r}
n = nrow(Ozone)
p = 4
X = cbind(rep(1,n), Ozone$wind, Ozone$humidity, Ozone$temp)
y = Ozone$ozone

beta_hat_no_lm = as.vector(solve(t(X) %*% X) %*% t(X) %*% y)
beta_hat_no_lm
sum(beta_hat_no_lm ^ 2)
```

**(b)** Obtain the estimated regression coefficients **with** the use of `lm()`. Store the results in a vector `beta_hat_lm`. To ensure this is a vector, you may need to use `as.vector()`. Return this vector as well as the results of `sum(beta_hat_lm ^ 2)`.

```{r}
ozone_model = lm(ozone ~ wind + humidity + temp, data = Ozone)
beta_hat_lm = as.vector(coef(ozone_model))
beta_hat_lm
sum(beta_hat_lm ^ 2)
```

**(c)** Use the `all.equal()` function to verify that the results are the same. You may need to remove the names of one of the vectors. The `as.vector()` function will do this as a side effect, or you can directly use `unname()`.

```{r}
all.equal(beta_hat_no_lm, beta_hat_lm)
```

**(d)** Calculate $s_e$ without the use of `lm()`. That is, continue with your results from **(a)** and perform additional matrix operations to obtain the result. Output this result. Also, verify that this result is the same as the result obtained from `lm()`.

```{r}
y_hat = X %*% solve(t(X) %*% X) %*% t(X) %*% y
s_e_no_lm = sqrt(sum((y - y_hat) ^ 2) / (n - p))
s_e_no_lm
s_e_lm = summary(ozone_model)$sigma
s_e_lm
all.equal(s_e_no_lm, s_e_lm)
```

**(e)** Calculate $R^2$ without the use of `lm()`. That is, continue with your results from **(a)** and **(d)**, and perform additional operations to obtain the result. Output this result. Also, verify that this result is the same as the result obtained from `lm()`.

```{r}
y_bar = mean(y)
r_2_no_lm = 1 - sum((y - y_hat) ^ 2) / sum((y - y_bar) ^ 2)
r_2_no_lm
r_2_lm = summary(ozone_model)$r.squared
r_2_lm
all.equal(r_2_no_lm, r_2_lm)
```

***

## Exercise 4 (Regression for Prediction)

For this exercise use the `Auto` dataset from the `ISLR` package. Use `?Auto` to learn about the dataset. The goal of this exercise is to find a model that is useful for **predicting** the response `mpg`. We remove the `name` variable as it is not useful for this analysis. (Also, this is an easier to load version of data from the textbook.)

```{r}
# load required package, remove "name" variable
library(ISLR)
Auto = subset(Auto, select = -c(name))
```

When evaluating a model for prediction, we often look at RMSE. However, if we both fit the model with all the data as well as evaluate RMSE using all the data, we're essentially cheating. We'd like to use RMSE as a measure of how well the model will predict on *unseen* data. If you haven't already noticed, the way we had been using RMSE resulted in RMSE decreasing as models became larger.

To correct for this, we will only use a portion of the data to fit the model, and then we will use leftover data to evaluate the model. We will call these datasets **train** (for fitting) and **test** (for evaluating). The definition of RMSE will stay the same

\[
\text{RMSE}(\text{model, data}) = \sqrt{\frac{1}{n} \sum_{i = 1}^{n}(y_i - \hat{y}_i)^2}
\]

where

- $y_i$ are the actual values of the response for the given data.
- $\hat{y}_i$ are the predicted values using the fitted model and the predictors from the data.

However, we will now evaluate it on both the **train** set and the **test** set separately. So each model you fit will have a **train** RMSE and a **test** RMSE. When calculating **test** RMSE, the predicted values will be found by predicting the response using the **test** data with the model fit using the **train** data. *__Test__ data should never be used to fit a model.*

- Train RMSE: Model fit with *train* data. Evaluate on **train** data.
- Test RMSE: Model fit with *train* data. Evaluate on **test** data.

Set a seed of `11`, and then split the `Auto` data into two datasets, one called `auto_trn` and one called `auto_tst`. The `auto_trn` data frame should contain 292 randomly chosen observations. The `auto_tst` data will contain the remaining observations. Hint: consider the following code:

```{r}
set.seed(11)
auto_trn_idx = sample(1:nrow(Auto), 292)
auto_trn = Auto[auto_trn_idx, ]
auto_tst = Auto[-auto_trn_idx, ]
```

Fit a total of five models using the training data.

- One must use all possible predictors.
- One must use only `displacement` as a predictor.
- The remaining three you can pick to be anything you like. One of these should be the *best* of the five for predicting the response.

For each model report the **train** and **test** RMSE. Arrange your results in a well-formatted markdown table. Argue that one of your models is the best for predicting the response.

```{r}
# model 1: all predictors
auto_model_1 = lm(mpg ~ ., data = auto_trn)
e = summary(auto_model_1)$residuals
RMSE_trn = sqrt(sum(e ^ 2) / length(e))
RMSE_trn
y_hat = predict(auto_model_1, newdata = auto_tst)
y = auto_tst$mpg
RMSE_tst = sqrt(sum((y - y_hat) ^ 2) / length(y))
RMSE_tst
```

```{r}
# model 2: only displacement
auto_model_2 = lm(mpg ~ displacement, data = auto_trn)
e = summary(auto_model_2)$residuals
RMSE_trn = sqrt(sum(e ^ 2) / length(e))
RMSE_trn
y_hat = predict(auto_model_2, newdata = auto_tst)
y = auto_tst$mpg
RMSE_tst = sqrt(sum((y - y_hat) ^ 2) / length(y))
RMSE_tst
```

```{r}
# model 3: displacement, weight, year, origin, cylinders
auto_model_3 = lm(mpg ~ displacement + weight + year + origin + cylinders, data = auto_trn)
e = summary(auto_model_3)$residuals
RMSE_trn = sqrt(sum(e ^ 2) / length(e))
RMSE_trn
y_hat = predict(auto_model_3, newdata = auto_tst)
y = auto_tst$mpg
RMSE_tst = sqrt(sum((y - y_hat) ^ 2) / length(y))
RMSE_tst
```

```{r}
# model 4: displacement, weight, year, origin
auto_model_4 = lm(mpg ~ displacement + weight + year + origin, data = auto_trn)
e = summary(auto_model_4)$residuals
RMSE_trn = sqrt(sum(e ^ 2) / length(e))
RMSE_trn
y_hat = predict(auto_model_4, newdata = auto_tst)
y = auto_tst$mpg
RMSE_tst = sqrt(sum((y - y_hat) ^ 2) / length(y))
RMSE_tst
```

```{r}
# model 5: weight, year, origin
auto_model_5 = lm(mpg ~ weight + year + origin, data = auto_trn)
e = summary(auto_model_5)$residuals
RMSE_trn = sqrt(sum(e ^ 2) / length(e))
RMSE_trn
y_hat = predict(auto_model_5, newdata = auto_tst)
y = auto_tst$mpg
RMSE_tst = sqrt(sum((y - y_hat) ^ 2) / length(y))
RMSE_tst
```

|  model  | RMSE_trn | RMSE_tst |
|:-------:|:--------:|:--------:|
| model 1 |  3.22365 | 3.572856 |
| model 2 | 4.553319 | 4.853014 |
| model 3 | 3.245324 |  3.56625 |
| model 4 | 3.264123 | 3.533147 |
| model 5 | 3.274669 | 3.518997 |

Since the model 5 has the smallest RMSE_tst, it is the best model for prediction with predictors of weight, year, and origin. 


***

## Exercise 5 (Simulating Multiple Regression)

For this exercise we will simulate data from the following model:

\[
Y_i = \beta_0 + \beta_1 x_{i1} + \beta_2 x_{i2} + \beta_3 x_{i3} + \beta_4 x_{i4} + \beta_5 x_{i5} + \epsilon_i
\]

Where $\epsilon_i \sim N(0, \sigma^2).$ Also, the parameters are known to be:

- $\beta_0 = 2$
- $\beta_1 = -0.75$
- $\beta_2 = 1.5$
- $\beta_3 = 0$
- $\beta_4 = 0$
- $\beta_5 = 2$
- $\sigma^2 = 25$

We will use samples of size `n = 42`.

We will verify the distribution of $\hat{\beta}_2$ as well as investigate some hypothesis tests.

**(a)** We will first generate the $X$ matrix and data frame that will be used throughout the exercise. Create the following nine variables:

- `x0`: a vector of length `n` that contains all `1`
- `x1`: a vector of length `n` that is randomly drawn from a normal distribution with a mean of `0` and a standard deviation of `2`
- `x2`: a vector of length `n` that is randomly drawn from a uniform distribution between `0` and `4`
- `x3`: a vector of length `n` that is randomly drawn from a normal distribution with a mean of `0` and a standard deviation of `1`
- `x4`: a vector of length `n` that is randomly drawn from a uniform distribution between `-2` and `2`
- `x5`: a vector of length `n` that is randomly drawn from a normal distribution with a mean of `0` and a standard deviation of `2`
- `X`: a matrix that contains `x0`, `x1`, `x2`, `x3`, `x4`, and `x5` as its columns
- `C`: the $C$ matrix that is defined as $(X^\top X)^{-1}$
- `y`: a vector of length `n` that contains all `0`
- `sim_data`: a data frame that stores `y` and the **five** *predictor* variables. `y` is currently a placeholder that we will update during the simulation.

Report the sum of the diagonal of `C` as well as the 5th row of `sim_data`. For this exercise we will use the seed `420`. Generate the above variables in the order listed after running the code below to set a seed.

```{r}
set.seed(420)
sample_size = 42
x0 = rep(1, n)
x1 = rnorm(n, mean = 0, sd = 2)
x2 = runif(n, min = 0, max = 4)
x3 = rnorm(n, mean = 0, sd = 1)
x4 = runif(n, min = -2, max = 2)
x5 = rnorm(n, mean = 0, sd = 2)
X = cbind(x0,x1,x2,x3,x4,x5)
C = solve(t(X) %*% X)
y = rep(0, n)
sim_data = data.frame(y,x1,x2,x3,x4,x5)
sum(diag(C))
sim_data[5, ]
```

**(b)** Create three vectors of length `2500` that will store results from the simulation in part **(c)**. Call them `beta_hat_1`, `beta_3_pval`, and `beta_5_pval`.

```{r}
num_sims = 2500
beta_hat_1 = rep(0,num_sims)
beta_3_pval = rep(0, num_sims)
beta_5_pval = rep(0, num_sims)
```

**(c)** Simulate 2500 samples of size `n = 42` from the model above. Each time update the `y` value of `sim_data`. Then use `lm()` to fit a multiple regression model. Each time store:

- The value of $\hat{\beta}_1$ in `beta_hat_1`
- The p-value for the two-sided test of $\beta_3 = 0$ in `beta_3_pval`
- The p-value for the two-sided test of $\beta_5 = 0$ in `beta_5_pval`

```{r}
beta_0 = 2
beta_1 = -0.75
beta_2 = 1.5
beta_3 = 0
beta_4 = 0
beta_5 = 2
sigma = 5

for (i in 1:num_sims) {
  eps = rnorm(n, mean = 0, sd = sigma)
  sim_data$y = beta_0 + beta_1 * x1 + beta_2 * x2 + beta_3 * x3 + beta_4 * x4 + beta_5 * x5 + eps
  fit_model = lm(y ~ ., data = sim_data)
  beta_hat_1[i] = coef(fit_model)[2]
  beta_3_pval[i] = summary(fit_model)$coefficients[4,4]
  beta_5_pval[i] = summary(fit_model)$coefficients[6,4]
}
```

**(d)** Based on the known values of $X$, what is the true distribution of $\hat{\beta}_1$?

```{r}
C_11 = C[2,2]
sigma^2 * C_11
```

$\hat{\beta}_1$ follows a normal distribution, $\hat\beta_1$ ~ $N(\beta_1, \sigma ^ 2 * C_{11})$.
That is, $\hat\beta_1$ ~ $N(-0.75, 0.01995094)$.

**(e)** Calculate the mean and variance of `beta_hat_1`. Are they close to what we would expect? Plot a histogram of `beta_hat_1`. Add a curve for the true distribution of $\hat{\beta}_1$. Does the curve seem to match the histogram?

```{r}
mean(beta_hat_1)
var(beta_hat_1)
```

The mean of `beta_hat_1` is -0.7561263, which is close to -0.75.
The variance of `beta_hat_1` is 0.01868924, which is close to 0.01995094.

```{r}
hist(beta_hat_1, prob = TRUE, breaks = 20, 
     xlab = expression(hat(beta)[1]), main = "", border = "dodgerblue")
curve(dnorm(x, mean = beta_1, sd = sqrt(sigma ^ 2 * C[2, 2])), 
      col = "darkorange", add = TRUE, lwd = 3)
```

The curve matches the histogram well. 

**(f)** What proportion of the p-values stored in `beta_3_pval` is less than 0.10? Is this what you would expect?

```{r}
mean(beta_3_pval < 0.10)
```

Since $\beta_3 = 0$, we expect 10% of the p_values to be significant at $\alpha = 0.1$ by chance, so it matches the expectation.  

**(g)** What proportion of the p-values stored in `beta_5_pval` is less than 0.01? Is this what you would expect?

```{r}
mean(beta_5_pval < 0.01)
```

Since $\beta_5 \neq 0$, we expect most of the p_values to be significant at $\alpha = 0.01$, so it matches the expectation.