---
title: "Week 3 - Homework"
author: "STAT 420, Summer 2020, Chenyue Jiao (cjiao4)"
date: 'June 6, 2020'
output:
  html_document: 
    theme: readable
    toc: yes  
  pdf_document: default
urlcolor: cyan
---


***

## Exercise 1 (Using `lm` for Inference)

For this exercise we will use the `cats` dataset from the `MASS` package. You should use `?cats` to learn about the background of this dataset.

**(a)** Fit the following simple linear regression model in `R`. Use heart weight as the response and body weight as the predictor. 

\[
Y_i = \beta_0 + \beta_1 x_i + \epsilon_i
\]

Store the results in a variable called `cat_model`. Use a $t$ test to test the significance of the regression. Report the following:

- The null and alternative hypotheses
- The value of the test statistic
- The p-value of the test
- A statistical decision at $\alpha = 0.05$
- A conclusion in the context of the problem

When reporting these, you should explicitly state them in your document, not assume that a reader will find and interpret them from a large block of `R` output.

```{r}
library(MASS)
cat_model = lm(Hwt ~ Bwt, data = cats)
summary(cat_model)
summary(cat_model)$coefficients
```

- The null and alternative hypotheses：
  The null hypothesis is $H_0: \beta_1 = 0$；
  The alternative hypothesis is $H_1: \beta_1 \neq 0$

- The value of the test statistic is 16.1193908 

- The p-value of the test is 6.969045e-34

- A statistical decision at $\alpha = 0.05$：
  We would reject the null hypothesis ($H_0: \beta_1 = 0$) at $\alpha = 0.05$ with the extremely low p-value. 

- A conclusion in the context of the problem：
  There is a significance liner relationship between the heart and body weights.


**(b)** Calculate a 95% confidence interval for $\beta_1$. Give an interpretation of the interval in the context of the problem.

```{r}
confint(cat_model, level = 0.95)[2, ]
```

The 95% confidence intervals for $\beta_1$ is between 3.539343 and 4.528782. We are 95% confident that for an increase of 1 kg in body weight, the average increase in heart weight is between 3.539343 and 4.528782 g.

**(c)** Calculate a 90% confidence interval for $\beta_0$. Give an interpretation of the interval in the context of the problem.

```{r}
confint(cat_model, level = 0.90)[1, ]
```

The 90% confidence intervals for $\beta_0$ is between -1.5028345 and 0.7895096. We are 90% confident that the average heart weight of a cat with 0 kg body weight is between -1.5028345 and 0.7895096 g, but we don’t believe that, since it would not be negative.


**(d)** Use a 90% confidence interval to estimate the mean heart weight for body weights of 2.1 and 2.8 kilograms. Which of the two intervals is wider? Why?

```{r}
predict(cat_model, newdata = data.frame(Bwt = c(2.1, 2.8)), 
        interval = c("confidence"), level = 0.90)
mean(cats$Bwt)
```

The 90% confidence interval to estimate the mean heart weight for body weight of 2.1 kg is between 7.787882 and 8.441856. The 90% confidence interval to estimate the mean heart weight for body weight of 2.8 kg is between 10.735843 and 11.141583. The interval of body weight of 2.1 kg is wider because the $x_i$ (2.1 kg) is far from the mean of $x$ (2.723611 kg), so the $(x - mean(x)) ^ 2$ is larger and thus the interval is wider.


**(e)** Use a 90% prediction interval to predict the heart weight for body weights of 2.8 and 4.2 kilograms.

```{r}
predict(cat_model, newdata = data.frame(Bwt = c(2.8, 4.2)), 
        interval = c("prediction"), level = 0.90)
```

The 90% prediction interval to predict the heart weight for body weight of 2.8 kg is between 8.525541 and 13.35189. The 90% prediction interval to predict the heart weight for body weight of 4.2 kg is between 14.097100 and 19.07570.


**(f)** Create a scatterplot of the data. Add the regression line, 95% confidence bands, and 95% prediction bands.

```{r}
Bwt_grid = seq(min(cats$Bwt), max(cats$Bwt), by = 0.01)
Hwt_ci_band = predict(cat_model, newdata = data.frame(Bwt = Bwt_grid), interval = "confidence", level = 0.95)
Hwt_pi_band = predict(cat_model, newdata = data.frame(Bwt = Bwt_grid), interval = "prediction", level = 0.95)

plot(Hwt ~ Bwt, data = cats,
     xlab = "Body Weight (in kg)",
     ylab = "Heart Weight (in g)",
     main = "Heart vs Body Weight",
     pch = 20,
     cex = 2,
     col = "grey",
     ylim = c(min(Hwt_pi_band), max(Hwt_pi_band)))
abline(cat_model, lwd = 5, col = "darkorange")

lines(Bwt_grid, Hwt_ci_band[,"lwr"], col = "dodgerblue", lwd = 3, lty = 2)
lines(Bwt_grid, Hwt_ci_band[,"upr"], col = "dodgerblue", lwd = 3, lty = 2)
lines(Bwt_grid, Hwt_pi_band[,"lwr"], col = "dodgerblue", lwd = 3, lty = 3)
lines(Bwt_grid, Hwt_pi_band[,"upr"], col = "dodgerblue", lwd = 3, lty = 3)
```


**(g)** Use a $t$ test to test:

- $H_0: \beta_1 = 4$
- $H_1: \beta_1 \neq 4$

Report the following:

- The value of the test statistic
- The p-value of the test
- A statistical decision at $\alpha = 0.05$

When reporting these, you should explicitly state them in your document, not assume that a reader will find and interpret them from a large block of `R` output.

```{r}
x = cats$Bwt
e = cat_model$residuals
beta_1_hat = coef(cat_model)[2]
Sxx = sum((x - mean(x)) ^ 2)
Se = sqrt(sum(e ^ 2) / (length(e) - 2))
t = (beta_1_hat - 4) * sqrt(Sxx) / Se
p_value = 2 * pt(abs(t), df = length(e) - 2, lower.tail = FALSE)
t
p_value
```

- The value of the test statistic is 0.1361084
- The p-value of the test is 0.8919283
- A statistical decision at $\alpha = 0.05$:
  We would **fail to reject** the null hypothesis ($H_0: \beta_1 = 4$) at $\alpha = 0.05$ with the large p-value.


***

## Exercise 2 (More `lm` for Inference)

For this exercise we will use the `Ozone` dataset from the `mlbench` package. You should use `?Ozone` to learn about the background of this dataset. You may need to install the `mlbench` package. If you do so, do not include code to install the package in your `R` Markdown document.

For simplicity, we will re-perform the data cleaning done in the previous homework.

```{r eval=FALSE, echo=FALSE}
install.packages("mlbench")
library(mlbench)
```

```{r}
data(Ozone, package = "mlbench")
Ozone = Ozone[, c(4, 6, 7, 8)]
colnames(Ozone) = c("ozone", "wind", "humidity", "temp")
Ozone = Ozone[complete.cases(Ozone), ]
```

**(a)** Fit the following simple linear regression model in `R`. Use the ozone measurement as the response and wind speed as the predictor. 

\[
Y_i = \beta_0 + \beta_1 x_i + \epsilon_i
\]

Store the results in a variable called `ozone_wind_model`. Use a $t$ test to test the significance of the regression. Report the following:

- The null and alternative hypotheses
- The value of the test statistic
- The p-value of the test
- A statistical decision at $\alpha = 0.01$
- A conclusion in the context of the problem

When reporting these, you should explicitly state them in your document, not assume that a reader will find and interpret them from a large block of `R` output.

```{r}
ozone_wind_model = lm(ozone ~ wind, data = Ozone)
summary(ozone_wind_model)
```

- The null and alternative hypotheses:
  The null hypothesis is $H_0: \beta_1 = 0$;
  The alternative hypothesis is $H_1: \beta_1 \neq 0$
  
- The value of the test statistic is -0.219

- The p-value of the test is 0.827

- A statistical decision at $\alpha = 0.01$:
  We would **fail to reject** the null hypothesis ($H_0: \beta_1 = 0$) at $\alpha = 0.01$ with the large p-value 0.827.

- A conclusion in the context of the problem:
  There is not a significant linear relationship between ozone measurement and wind speed. 


**(b)** Fit the following simple linear regression model in `R`. Use the ozone measurement as the response and temperature as the predictor. 

\[
Y_i = \beta_0 + \beta_1 x_i + \epsilon_i
\]

Store the results in a variable called `ozone_temp_model`. Use a $t$ test to test the significance of the regression. Report the following:

- The null and alternative hypotheses
- The value of the test statistic
- The p-value of the test
- A statistical decision at $\alpha = 0.01$
- A conclusion in the context of the problem

When reporting these, you should explicitly state them in your document, not assume that a reader will find and interpret them from a large block of `R` output.

```{r}
ozone_temp_model = lm(ozone ~ temp, data = Ozone)
summary(ozone_temp_model)
summary(ozone_temp_model)$coefficients
```

- The null and alternative hypotheses:
  The null hypothesis is $H_0: \beta_1 = 0$;
  The alternative hypothesis is $H_1: \beta_1 \neq 0$
  
- The value of the test statistic is 22.85

- The p-value of the test is 8.153764e-71

- A statistical decision at $\alpha = 0.01$:
  We would reject the null hypothesis ($H_0: \beta_1 = 0$) at $\alpha = 0.01$ with the extremely low p-value. 

- A conclusion in the context of the problem:
  There is a significance liner relationship between ozone measurement and temperature.


***

## Exercise 3 (Simulating Sampling Distributions)

For this exercise we will simulate data from the following model:

\[
Y_i = \beta_0 + \beta_1 x_i + \epsilon_i
\]

Where $\epsilon_i \sim N(0, \sigma^2).$ Also, the parameters are known to be:

- $\beta_0 = -5$
- $\beta_1 = 3.25$
- $\sigma^2 = 16$

We will use samples of size $n = 50$.

**(a)** Simulate this model $2000$ times. Each time use `lm()` to fit a simple linear regression model, then store the value of $\hat{\beta}_0$ and $\hat{\beta}_1$. Set a seed using **your** birthday before performing the simulation. Note, we are simulating the $x$ values once, and then they remain fixed for the remainder of the exercise.

```{r}
birthday = 19950117
set.seed(birthday)
n = 50
x = seq(0, 10, length = n)
```

```{r}
beta_0 = -5
beta_1 = 3.25
sigma = 4

num_samples = 2000
beta_0_hats = rep(0, num_samples)
beta_1_hats = rep(0, num_samples)

for (i in 1:num_samples) {
  eps = rnorm(n, mean = 0, sd = sigma)
  y = beta_0 + beta_1 * x + eps
  
  sim_model = lm(y ~ x)
  
  beta_0_hats[i] = coef(sim_model)[1]
  beta_1_hats[i] = coef(sim_model)[2]
}
```


**(b)** Create a table that summarizes the results of the simulations. The table should have two columns, one for $\hat{\beta}_0$ and one for $\hat{\beta}_1$. The table should have four rows:

- A row for the true expected value given the known values of $x$
- A row for the mean of the simulated values
- A row for the true standard deviation given the known values of $x$
- A row for the standard deviation of the simulated values

```{r}
mean(beta_0_hats)
mean(beta_1_hats)

Sxx = sum((x - mean(x)) ^2)
sigma * sqrt(1 / n + mean(x) ^ 2 / Sxx)
sigma / sqrt(Sxx)

sd(beta_0_hats)
sd(beta_1_hats)
```
|                | $\hat{\beta}_0$ | $\hat{\beta}_1$ |
|:--------------:|:---------------:|:---------------:|
|  true_expected |        -5       |       3.25      |
| simulated_mean |    -5.012482    |     3.248514    |
|     true_sd    |     1.114609    |    0.1920784    |
|  simulated_sd  |     1.135973    |    0.1965515    |


**(c)** Plot two histograms side-by-side:

- A histogram of your simulated values for $\hat{\beta}_0$. Add the normal curve for the true sampling distribution of $\hat{\beta}_0$.
- A histogram of your simulated values for $\hat{\beta}_1$. Add the normal curve for the true sampling distribution of $\hat{\beta}_1$.

```{r}
par(mfrow=c(1,2))
hist(beta_0_hats, prob = TRUE, breaks = 25, 
     xlab = expression(hat(beta)[0]), main = "", border = "dodgerblue")
curve(dnorm(x, mean = -5, sd = 1.114609),
      col = "darkorange", add = TRUE, lwd = 3)

hist(beta_1_hats, prob = TRUE, breaks = 25,
     xlab = expression(hat(beta)[1]), main = "", border = "dodgerblue")
curve(dnorm(x, mean = 3.25, sd = 0.1920784), col = "darkorange", add = TRUE, lwd = 3)
```


***

## Exercise 4 (Simulating Confidence Intervals)

For this exercise we will simulate data from the following model:

\[
Y_i = \beta_0 + \beta_1 x_i + \epsilon_i
\]

Where $\epsilon_i \sim N(0, \sigma^2).$ Also, the parameters are known to be:

- $\beta_0 = 5$
- $\beta_1 = 2$
- $\sigma^2 = 9$

We will use samples of size $n = 25$.

Our goal here is to use simulation to verify that the confidence intervals really do have their stated confidence level. Do **not** use the `confint()` function for this entire exercise.

**(a)** Simulate this model $2500$ times. Each time use `lm()` to fit a simple linear regression model, then store the value of $\hat{\beta}_1$ and $s_e$. Set a seed using **your** birthday before performing the simulation. Note, we are simulating the $x$ values once, and then they remain fixed for the remainder of the exercise.

```{r}
birthday = 19950117
set.seed(birthday)
n = 25
x = seq(0, 2.5, length = n)
```

```{r}
beta_0 = 5
beta_1 = 2
sigma = 3

num_samples = 2500
beta_1_hats = rep(0, num_samples)
beta_1_hat_se = rep(0, num_samples)

for (i in 1:num_samples) {
  eps = rnorm(n, mean = 0, sd = sigma)
  y = beta_0 + beta_1 * x + eps
  
  sim_model = lm(y ~ x)
  e = sim_model$residuals
  beta_1_hats[i] = coef(sim_model)[2]
  beta_1_hat_se[i] = summary(sim_model)$coefficients[2,2]
}
```


**(b)** For each of the $\hat{\beta}_1$ that you simulated, calculate a 95% confidence interval. Store the lower limits in a vector `lower_95` and the upper limits in a vector `upper_95`. Some hints:

- You will need to use `qt()` to calculate the critical value, which will be the same for each interval.
- Remember that `x` is fixed, so $S_{xx}$ will be the same for each interval.
- You could, but do not need to write a `for` loop. Remember vectorized operations.

```{r}
crit = qt(0.975, df = num_samples - 2)
lower_95 = beta_1_hats - crit * beta_1_hat_se
upper_95 = beta_1_hats + crit * beta_1_hat_se
```


**(c)** What proportion of these intervals contains the true value of $\beta_1$?

```{r}
mean(lower_95 < 2 & 2 < upper_95)
```


**(d)** Based on these intervals, what proportion of the simulations would reject the test $H_0: \beta_1 = 0$ vs $H_1: \beta_1 \neq 0$ at $\alpha = 0.05$?

```{r}
1 - mean(lower_95 < 0 & 0 < upper_95)
```


**(e)** For each of the $\hat{\beta}_1$ that you simulated, calculate a 99% confidence interval. Store the lower limits in a vector `lower_99` and the upper limits in a vector `upper_99`.

```{r}
crit = qt(0.995, df = num_samples - 2)
lower_99 = beta_1_hats - crit * beta_1_hat_se
upper_99 = beta_1_hats + crit * beta_1_hat_se
```

**(f)** What proportion of these intervals contains the true value of $\beta_1$?

```{r}
mean(lower_99 < 2 & 2 < upper_99)
```

**(g)** Based on these intervals, what proportion of the simulations would reject the test $H_0: \beta_1 = 0$ vs $H_1: \beta_1 \neq 0$ at $\alpha = 0.01$?

```{r}
1 - mean(lower_99 < 0 & 0 < upper_99)
```


***

## Exercise 5 (Prediction Intervals "without" `predict`)

Write a function named `calc_pred_int` that performs calculates prediction intervals:

$$
\hat{y}(x) \pm t_{\alpha/2, n - 2} \cdot s_e\sqrt{1 + \frac{1}{n}+\frac{(x-\bar{x})^2}{S_{xx}}}.
$$

for the linear model

$$
Y_i = \beta_0 + \beta_1 x_i + \epsilon_i.
$$

**(a)** Write this function. You may use the `predict()` function, but you may **not** supply a value for the `level` argument of `predict()`. (You can certainly use `predict()` any way you would like in order to check your work.)

The function should take three inputs:

- `model`, a model object that is the result of fitting the SLR model with `lm()`
- `newdata`, a data frame with a single observation (row)
    - This data frame will need to have a variable (column) with the same name as the data used to fit `model`.
- `level`, the level (0.90, 0.95, etc) for the interval with a default value of `0.95`

The function should return a named vector with three elements:

- `estimate`, the midpoint of the interval
- `lower`, the lower bound of the interval
- `upper`, the upper bound of the interval

```{r}
calc_pred_int = function(model, newdata, level = 0.95) {
  a = 1 - level
  n = length(resid(model))
  beta_0_hat = coef(model)[1]
  beta_1_hat = coef(model)[2]
  y_hat = fitted(model)
  x = (y_hat - beta_0_hat) / beta_1_hat
  
  s_e = summary(model)$sigma
  Sxx = sum((x - mean(x)) ^2)
  
  crit = qt((1 - a / 2), df = n - 2)
  estimate = beta_0_hat + beta_1_hat * newdata
  lower = estimate - crit * s_e * sqrt(1 + 1 / n + (newdata - mean(x)) ^ 2 / Sxx)
  upper = estimate + crit * s_e * sqrt(1 + 1 / n + (newdata - mean(x)) ^ 2 / Sxx)
  
  c("estimate" = as.numeric(estimate), "lower" = as.numeric(lower), "upper" = as.numeric(upper))
}
```


**(b)** After writing the function, run this code:

```{r}
newcat_1 = data.frame(Bwt = 4.0)
calc_pred_int(cat_model, newcat_1)
```


**(c)** After writing the function, run this code:

```{r}
newcat_2 = data.frame(Bwt = 3.3)
calc_pred_int(cat_model, newcat_2, level = 0.90)
```

